\chapter{INTRODUCTION}
\thispagestyle{empty}

\mquote{The Universe is under no obligation to make sense to you.}{Neil deGrasse Tyson}

Accretion of various types is one of the prevalent mechanisms by which matter is distributed and mixed in the universe. The primary focus of this study is computer modeling of accretion disks, and in particular, accretion disks of Cataclysmic Variable stars.

% kapitola 2
In Chapter \ref{chap:accretion_disks}, we start by defining what we mean by the term accretion and the different types of accreting systems, their basic characteristics, and a few notable examples of such systems. We then discuss different types of accretion disk models in relation to the \emph{Eddington luminosity limit}, and an extra focus is given to the Shakura-Sunyaev $\alpha$-disk model, which is also referred to as the \emph{stardard thin model}. It is the $\alpha$-disk model whose parameters, and primarily the free parameter $\alpha$, that we aim to obtain by utilizing our models discussed in the latter chapters.

% kapitola 5
Chapter \ref{chap:dripping_faucet} provides an overview of different types of dripping faucets models and the theory behind them. There are also simulation results examples attached that were made by the author and are based on the discussed theory and previous work by \citep{kveton2014}.

% kapitola 6
Chapter \ref{chap:multilayer_dripping_handrail} focuses primarily on the definition of our \emph{Multilayer Dripping Handrail} model (MDH) and its parameters. This Chapter also aims to provide precise explanations of algorithms and processes used to simulate this system and methods by which we extract synthetic observations from the dynamical matter flow simulation results.

% kapitola 7
In Chapter \ref{chap:model_implementation}, we delve deeper into the technical side of the MDH model, discuss the actual code implementation, and explain the concepts and technologies used to construct the simulation code. Also, we discussed some hardware and software limitations and obstacles we faced. This Chapter should serve as an overview of our code and maybe an inspiration for similar experiments in the future.

% kapitola 8
The examples of MDH results are presented in Chapter~\ref{chap:cv_models}, where we demonstrate the behavior and customizability of MDH results by computing several edge cases with combinations of high and low values of MDH's free parameters. 

% kapitola 9
In Chapter~\ref{chap:alpha_model_fitting}, we used the same MDH results to compute the mean area density distribution of the simulated accretion disk, which we fitted with the Shakura-Sunyaev $\alpha$-disk model solution. The analytical solution of Shakura-Sunyaev's model agrees closely with our purely numerical approach. Therefore we can obtain a very accurate value of the free parameter $\alpha$.
